### 2013-11-07 paper/inference.tex
major changes to formulation part

\subsection{Bayesian Inference}
 \label{sec:inference}
Bayesian inference is a statistical approach to inverse problems
that has recently gained much interest in different applications.
The key idea in this approach is to express all forms of uncertainty
in terms of probability. The first step in a Bayesian inference 
is to formulate the forward problem (response surface/model) using 
a suitable likelihood function and product of conditional probability densities. 
We then formulate the prior distribution over the unknown parameters of the model
using our best belief. Lastly, given some observation data, Bayes rule 
is used to obtain a posterior distribution for these unknown
\citep{sivia}. We review this approach briefly below.

Let the observation data (water surface elevation at the gauges) be denoted by \{$S_i = S_i(x_i,y_i,t_i)\}_{i=1}^m$;
their model predicted counterparts by \{$M_i= M_i(x_i,y_i,t_i)\}_{i=1}^m$
($m$ is the number of observations) and $\{N_i\}_{i=1}^3$ the set of 
three Manning Friction coefficients treated as the unknown parameters. 
Bayes' rule yields:
\begin{equation}
 p(\{N_i\}_{i=1}^3| \{S_i\}_{i=1}^m) \propto 
 p(\{S_i\}_{i=1}^m | \{N_i\}_{i=1}^3) \ p(\{N_i\}_{i=1}^3)  
\label{eq:bayes}
\end{equation}
where $p(\{N_i\}_{i=1}^3)$ is the prior of $\{N_i\}_{i=1}^3$, representing the \emph{a priori} knowledge
about the manning coefficient; 
$p(\{S_i\}_{i=1}^m | \{N_i\}_{i=1}^3)$ is the likelihood function representing
the probability of obtaining the data given the set of parameters $\{N_i\}_{i=1}^3$;
and finally $p(\{N_i\}_{i=1}^3| \{ S_i \}_{i=1}^m)$ is the posterior,
representing the probability that $\{N_i\}_{i=1}^3$ is true given the data $( \{ S_i \}_{i=1}^m )$.

\begin{equation}
p(\{N_i\}_{i=1}^3 | \{ S_i \}_{i=1}^m)
\propto 
p(\{S_i\}_{i=1}^m | \{N_i\}_{i=1}^3) \ p(N_1)p(N_2) p(N_3)  
\end{equation}

To formulate the likelihood function, we let 
	 $\epsilon_i = S_i - M_i $
represents the discrepancy between the model and observations 
and assume that the observations are independent
and normally distributed with mean
zero and variance $\sigma^2$, i.e. 
$\epsilon_i \sim N(0,\sigma^2)$. In this 
case the likelihood function can be written as:
\begin{equation} 
p(\{S_i\}_{i=1}^m |  \{N_i\}_{i=1}^3) 
= 
\prod_{i=1}^m  \frac{1}{\sqrt{2 \pi \sigma^2}} 
\exp \left( \frac{-(S_i - M_i)^2}{2 \sigma^2} \right)  	
\label{eq:likelihood}
\end{equation}

The variance $\sigma^2$ is unknown \emph{a priori}; thus we treat it as a hyper-parameter.
While in general $\sigma^2$ depends on the observations, in cases where the error
amplitude is generally small and does not change throughout space and time, one may use
a simplified model and assume single hyper-parameter $\sigma^2$.

The next step is choosing the prior that should be based 
on some \emph{a priori} knowledge about the parameters. In our case, a uniform
prior for the model parameters is assumed:
\begin{equation} 
p(N_i) = \begin{cases}
		\displaystyle \frac{1}{b_i-a_i} &\text{for~} a_i <  N_i \leq b_i ,  \\
		0 &\text{otherwise}  , 
\end{cases}
\end{equation}
where $ [a_i,b_i]$ denote the parameter ranges defined in 
Equations~(\ref{eq:prior1}--\ref{eq:prior3}).
Regarding the variance, the only information we know 
is that $\sigma^2$ is always positive.
We thus assume a Jeffreys prior \citep{sivia}, expressed as:
\begin{equation} 
p(\sigma^2) =  \begin{cases}
		\displaystyle \frac{1}{\sigma^2} &\text{for~} \sigma^2 > 0,  \\
		0 &\text{otherwise}. 
		\end{cases}
\label{eq:var_pr}
\end{equation}
Consequently, Bayes' theorem gives:
\begin{equation} 
p(\{N_i\}_{i=1}^3,\sigma^2 | \{S_i\}_{i=1}^m) 
\propto
\left[ \prod_{i=1}^m  \frac{1}{\sqrt{2 \pi \sigma^2}} 
\exp \left( \frac{-(S_i - M_i)^2}{2 \sigma^2} \right) \right] 
\ p(\sigma^2)p(N_1)p(N_2) p(N_3)
\end{equation}
where $p(\{N_i\}_{i=1}^3,\sigma^2 | \{S_i\}_{i=1}^m)$  is the joint posterior.

Inferring the drag coefficient parameters requires 
sampling the posterior. In general, when the space of the unknown 
parameters is multidimensional, a suitable computational strategy is 
the Markov Chain Monte  Carlo (MCMC) method. 
We rely on an adaptive Metropolis MCMC \citep{Gareth2009,Haario2001} to
sample the posterior distribution accurately and efficiently.
This MCMC phase requires repeated (tens of
thousands of) Geoclaw simulations initialized with different values of the
uncertain parameters; this step is prohibitively expensive. An
alternative is to construct a surrogate model that requires a much
smaller ensemble of GeoClaw runs, and that can be used instead
at a significantly reduced computational cost.  Here, we rely on
PCEs to build the surrogate, which, in addition
also provide statistical properties,  such as the mean, variance and sensitivities,
efficiently.

>>>>

\subsection{Bayesian Inference}
 \label{sec:inference}

Bayesian inference is a statistical approach to inverse problems
that has recently gained much interest in different applications including
ocean~\citep{Alexanderian2011a,Zedler2012,sraj:2013a}
climate~\citep{OlsonEtAl2012} and geophysical~\citep{Malinverno2002} modeling.
The key idea in this approach is to express all forms of uncertainty
in terms of probability. The first step in a Bayesian inference 
is to formulate the forward problem (response surface/model) using 
a suitable likelihood function and a product of conditional probability densities. 
We then formulate the prior distribution over the unknown parameters of the model
using our best belief. Lastly, given some observation data, Bayes rule 
is used to obtain a posterior distribution for these unknown
\citep{sivia}. We review this approach briefly below.

Let $\vec{d}=(d_1,...,d_m)^T$ be a vector of observation data and $\vec{\theta}=(\theta_1,...,\theta_n)^T$ be a vector of model parameters or inputs. We now consider a forward model $\vec G$ that predicts the data function of 
the parameters such that:
\begin{equation}
\vec d \approx \vec{G}( \vec \theta)
\end{equation}

Applying the Bayes' rule yields:
\begin{equation}
 p(\vec{\theta}| \vec d) \propto 
 p(\vec d | \vec{\theta}) \ p(\vec{\theta})  
\label{eq:bayes}
\end{equation}
where $p(\vec{\theta})$ is the prior of $\vec{\theta}$, representing the \emph{a priori} knowledge
about the parameters; 
$p(\vec d| \vec{\theta})$ is the likelihood function representing
the probability of obtaining the data given the set of parameters $\vec{\theta}$;
and finally $p(\vec{\theta}| \vec d)$ is the posterior,
representing the probability that $\vec{\theta}$ is true given the data $\vec d $.

To formulate the likelihood function, we let $\vec \epsilon = \vec d - \vec{G}$
represents the discrepancy between the model and observations.
Here, the components of $\vec \epsilon $ are i.i.d. random variables with density $p_{\epsilon}$.
In this case the likelihood function can be written as:
\begin{equation} 
p(\vec d |  \vec{\theta}) 
= 
\prod_{i=1}^m  
p_\epsilon (d_i - G_i(\vec \theta))  	
\label{eq:likelihood}
\end{equation}

In our application, we assume that the errors $\epsilon_i$ are independent
and normally distributed with mean
zero and variance $\sigma^2$, i.e. 
$\epsilon_i \sim N(0,\sigma^2)$. 
While in general $\sigma^2$ depends on the observations, in cases where the error
amplitude is generally small and does not change throughout space and time, one may use
a simplified model and assume single $\sigma^2$ value.
Consequently, Equation~\ref{eq:likelihood} becomes:
\begin{equation} 
p(\vec d |  \vec{\theta}) 
= 
\frac{1}{\sqrt{2 \pi \sigma^2}}\prod_{i=1}^m   
\exp \left\lbrace \frac{-(d_i - G_i(\vec \theta))^2}{2 \sigma^2} \right\rbrace 	
\label{eq:likelihood2}
\end{equation}
The joint posterior from Bayes' rule can be given as:
\begin{equation} 
p(\vec{\theta}| \vec d)
\propto
\frac{1}{\sqrt{2 \pi \sigma^2}}   \prod_{i=1}^m  
\exp \left\lbrace \frac{-(d_i - G_i(\vec \theta))^2}{2 \sigma^2} \right\rbrace  
\prod_{i=1}^n p(\theta_i)
\end{equation}
The variance $\sigma^2$ is unknown \emph{a priori}; thus we treat it as a hyper-parameter.
In other words, $\sigma^2$ may become an additional object for Bayesian inference and may themselves be endowed with priors and estimated from data. In this case the joint posterior is finally written as follows:
\begin{equation} 
p(\vec{\theta},\sigma^2 | \vec d)
\propto
\frac{1}{\sqrt{2 \pi \sigma^2}}   \prod_{i=1}^m  
\exp \left\lbrace \frac{-(d_i - G_i(\vec \theta))^2}{2 \sigma^2} \right\rbrace
\ p(\sigma^2) \prod_{i=1}^n p(\theta_i)
\label{eq:post}
\end{equation}

### 2013-11-07 paper/mcmc.tex
major changes to formulation part

Before attempting solving the inverse problem,
we need to make sure that the observed data are
comparable to the simulated one.
To this end, Figure~\ref{fig:compare}
compares PC mean water surface elevation profiles
with their counterparts from the observed gauge locations.
The plots show good agreement between the simulations and the observations at different
locations and at different times. 
%
\begin{figure}[h]
\begin{tabular}{clc}
%        
\includegraphics[width=0.475\textwidth]{../figures/compare1.pdf} &
\includegraphics[width=0.475\textwidth]{../figures/compare2.pdf} \\
\includegraphics[width=0.475\textwidth]{../figures/compare3.pdf} &
\includegraphics[width=0.475\textwidth]{../figures/compare4.pdf}
\end{tabular}
\caption{Comparison of PC mean 
and observed data of water surface elevation with time at the four gauges.}
\label{fig:compare}
\end{figure}

Figure~\ref{fig:compare} can be recast as a scatter plot shown in 
Figure~\ref{fig:scatter}.
\begin{figure}[h]
\centering
\includegraphics[width=0.475\textwidth]{../figures/scatter.pdf} 
\caption{Scatter plot of PC mean water surface elevation vs. observed ones.}
\label{fig:scatter}

\end{figure}

@@

\begin{figure}[h]
\begin{tabular}{clc}
\includegraphics[width=0.475\textwidth]{../figures/chain_p1.pdf} &
\includegraphics[width=0.475\textwidth]{../figures/chain_p2.pdf} \\
\includegraphics[width=0.475\textwidth]{../figures/chain_p3.pdf} &
\includegraphics[width=0.475\textwidth]{../figures/chain_s1.pdf}
\end{tabular}
\caption{Chain samples for the three Manning's roughness coefficients $N_1,N_2,N_3$ and $\sigma^2$
the variance between simulations and observations.}
\label{fig:mcmc} 
\end{figure}

@@

 \begin{figure}[h]
        \begin{tabular}{clc}
\includegraphics[width=0.475\textwidth]{../figures/pdf_p1.pdf} &
\includegraphics[width=0.475\textwidth]{../figures/pdf_p2.pdf} \\
\includegraphics[width=0.475\textwidth]{../figures/pdf_p3.pdf} &
\includegraphics[width=0.475\textwidth]{../figures/pdf_s1.pdf}
        \end{tabular}
        \caption{Posterior distributions for the three Manning's roughness coefficients $N_1,N_2,N_3$ 
and $\sigma^2$ the variance between simulations and observations.}
\label{fig:pdfs} 
        \end{figure}

>>>>

Before attempting solving the inverse problem,
we need to make sure that the observed data are
comparable to the simulated one.
To this end, Figure~\ref{fig:compare}
compares PC mean water surface elevation profiles
with their counterparts from the observed gauge locations.
The plots show good agreement between the simulations and the observations at different
locations and at different times. 
%

Figure~\ref{fig:compare} can be recast as a scatter plot shown in 
Figure~\ref{fig:scatter}.

@@

@@

### 2013-11-07 paper/uqpce.tex
adding inference section
major changes to formulation part

\section{Polynomial Chaos Expansion}
\label{sec:uqpce}

Polynomial Chaos (PC) is a spectral method that have been developed in the engineering community to represent uncertainties in the output of numerical simulations \citep{Villegas2012,Lin2009,Xiu2004}
due to the uncertainties in a model's input.
The PC method expresses the dependencies of a quantity of 
interest on the uncertain input variables as a truncated expansion.

Let $U=U(\bm{x},t,\xxi)$ 
denote an output quantity of 
interest normally function of space $\bm{x}$ and time $t$, and also on the uncertain variables $\xxi$ that went into
specifying the model's input data (such as uncertain initial and boundary conditions, and parameters). PC expansion express $U$ in the form:
\begin{equation}
  U(\xbold,t,\xxi) \doteq \sum_{k = 0}^P U_k(\xbold,t) \Psi_k(\xxi)
\label{eq:stochseries}
\end{equation}
where $U_k(\xbold,t)$ are the polynomial coefficients, $P$ is finite
and depends on the truncation strategy adopted, and the functions
$\Psi_k(\xxi)$ form an orthogonal basis of an underlying probability
space. This series representation can be viewed as a spectral expansion
of $U$ along the stochastic dimensions.  The existence and convergence of this series
is asserted by the Cameron-Martin theorem
\citep{Cameron:1947}  whenever $R$ has finite variance. The series rate of convergence, and
hence the number of terms to retain, depends on the smoothness of
$U$ with respect to $\xxi$. The series converges spectrally fast with $P$
when $U$ is infinitely smooth; the convergence rate becomes algebraic
when $U$ has finite smoothness \citep{Canuto:2006}. 
%In practice the series convergence is monitored 
%via various error metrics; Section \ref{sec:results}\ref{sec:analysis} presents one possible error
%analysis for the present problem.

The choice of basis is dictated by the probability density
function of the stochastic variable $\xxi$, and which appears as a weight
function in the stochastic space's inner product:
\begin{equation}
 \left<\Psi_k,\Psi_m\right> = \int \Psi_k(\xxi) \;\Psi_m(\xxi) \; \rho(\xxi) \; \mbox{d}\xxi.
\end{equation}
When the basis functions are orthonormal $\left<\Psi_k,\Psi_m\right>=\delta_{k,m}$, where
$\delta_{k,m}$ is the Kronecker delta.
For uniform
distributions the basis functions are hence scaled Legendre polynomials.
For multi-dimensional problems the basis functions are
tensor products of 1D basis functions. The identification of the inner product weight function
with the probability distribution of $\xxi$ simplifies the calculations of $U$'s statistical moments.
Noting that since $\Psi_0(\xxi)$ is a constant that can be normalized to satisfy 
$\left<\Psi_0,\Psi_0\right>=1$, the expectation and variance of $U$ can be computed as:
\begin{eqnarray}
 E[U]&=&\int U \, \rho(\xxi) \, \mbox{d}\xxi=\left< U,\Psi_0\right> = U_0 \\
 E[(U-E[U])^2]&=&\int (U-E[U])^2 \, \rho(\xxi) \, \mbox{d}\xxi=\sum_{k=1}^P U_k^2
\left<\Psi_k,\Psi_k\right>
\end{eqnarray}
The series representation (\ref{eq:stochseries}) can thus be seen as
combining approximation and probabilistic frameworks, a combination
that has proven extremely useful in solving UQ
problems.

A number of procedures have been devised to compute the expansion coefficients,
$U_k$. Here we rely on the Non-Intrusive Spectral Projection (NISP) 
since it does not require any modification to the deterministic model, and only
ensemble runs at specified values of the uncertain parameters are needed.
The NISP method
exploits the orthogonality of the basis, and replaces the
stochastic integrals by quadrature to obtain:
\begin{equation}
 U_k(\bm{x},t) = \frac{\left< U, \Psi_k \right>}{\left< \Psi_k, \Psi_k \right>}
         \approx \frac{\left< U, \Psi_k \right>_Q}{\left< \Psi_k, \Psi_k \right>} .
\end{equation}
The subscript $Q$ refers to approximating the inner product integral with
quadrature, i.e.
\begin{equation}
  \left< U, \Psi_k \right> = \int U(\bm{x},t,\xxi) \Psi_k(\xxi) \rho(\xxi) \mbox{ d}\xxi
\approx \left< U, \Psi_k \right>_Q
= \sum_{q=1}^Q U(\xxi_q) \Psi_k(\xxi_q) \omega_q
\end{equation}
where $\xxi_q$ and $\omega_q$ are multi-dimensional quadrature points and weights,
respectively.
The computation of the ${U}_k$ can thus be expressed as a matrix-vector product of the form:
\begin{equation} 
 U_k(\bm{x},t)=\sum_q \Pi_{kq} U(\bm{x},t,\xxi_q),\;\;\;
 \Pi_{kq}=\frac{\Psi_k(\xxi_q)\omega_q}{\left< \Psi_k, \Psi_k \right>}
\end{equation} 
where $\Pi_{kq}$ is the projection matrix and $U(\bm{x},t,\xxi_q)$ is obtained
from an ensemble of the deterministic model {\em realizations} with the uncertain parameters set at
the quadrature value $\xxi_q$. The NISP procedure can be carried out as a post-processing
operation using simple scripting languages like MATLAB, the only costs being those of the
ensemble calculation and its storage.

The calculation of the $U(\bm{x},t,\xxi_q)$ is the most expensive part of the
inference procedure; thus, reducing the number of sampling points
while maximizing their effectiveness is critical for the procedure's
efficiency.  The quadrature order should be commensurate with the
truncation order, and should be high enough to avoid aliasing artifacts.
The choice of quadrature rule is hence critical to the performance
of the PC (in its NISP version at least). In this work, we employ the 
tensorized Gaussian
quadratures as their computational costs are affordable.
For more details on PC exapnsions and other UQ methods,
refer to \citep{LeMaitreKnio2010}.

>>>>

\subsection{Accelerating Bayesian Inference}
\label{sec:uqpce}

MCMC exploration of the 4D posterior (Equation~\ref{eq:post_coef}) requires repeated simulations (tens of thousands) of the forward GeoClaw model, once for each proposed move of the Markov chain. While a single \emph{Geoclaw} simulation
takes $\sim 15~$mins, depending on the details of the MCMC sampler, it is desirable to avoid repeated forward solutions altogether. An alternative is to construct a surrogate model that requires a much
smaller ensemble of GeoClaw runs, and that can be used instead
at a significantly reduced computational cost.  Here, we rely on
Polynomial Chaos expansions for accelerating Bayesian inference in this context 
by building a surrogate, which, in addition
also provide statistical properties,  such as the mean, variance and sensitivities,
efficiently.

\subsubsection{Polynomial Chaos}

Polynomial Chaos (PC) is a probabilistic methodology that expresses the 
dependencies of model outputs on the uncertain model inputs
as a polynomial truncated expansion. This method has been developed in 
the engineering community to represent uncertainties in the output of 
numerical simulations~\citep{Villegas2012,Lin2009,Xiu2004}
due to the uncertainties in a model's input. We briefly describe the PC
method below; for more details on this method, and other UQ methods
refer to \citep{LeMaitreKnio2010}.

Let $U=U(\bm{x},t,\xxi)$ denote a quantity of 
interest (QoI) that is the output of certain model.
$U$ is normally function of space $\bm{x}$ and time $t$, and 
also on the vector of normalized uncertain model inputs $\xxi=(\xi_1,...,\xi_n)$ (called the germ). PC expresses $U$ in the form:
\begin{equation}
  U(\xbold,t,\xxi) \doteq \sum_{k = 0}^P U_k(\xbold,t) \Psi_k(\xxi)
\label{eq:stochseries}
\end{equation} 
where $U_k(\xbold,t)$ are the polynomial coefficients, and
$\Psi_k(\xxi)$ are functions that form an orthogonal basis of an underlying probability
space. The total number of terms in the PC expansion is
$P+1 = \frac{(d+p)! }{n!\ p!}$ where $n$ is the number of stochastic dimension and $p$ is the highest order
of polynomials. 
This series representation can be viewed as a spectral expansion
of $U$ along the stochastic dimensions.  The existence and convergence of this series
is asserted by the Cameron-Martin theorem
\citep{Cameron:1947}  whenever $U$ has finite variance. The series rate of convergence, and
hence the number of terms to retain, depends on the smoothness of
$U$ with respect to $\xxi$. The series converges spectrally fast with $P$
when $U$ is infinitely smooth; the convergence rate becomes algebraic
when $U$ has finite smoothness \citep{Canuto:2006}. 
%In practice the series convergence is monitored 
%via various error metrics; Section \ref{sec:results}\ref{sec:analysis} presents one possible error
%analysis for the present problem.

The choice of basis is dictated by the probability density
function of the stochastic variable $\xxi$, and which appears as a weight
function in the stochastic space's inner product:
\begin{equation}
 \left<\Psi_i,\Psi_j\right> = \int \Psi_i(\xxi) \;\Psi_j(\xxi) \; \rho(\xxi) \; \mbox{d}\xxi=\delta_{ij}\ave{\Psi_i^2}
\label{eq:inner}
\end{equation}
where $\delta_{ij}$ is the Kronecker delta.
For uniform
distributions the basis functions are hence scaled Legendre polynomials.
For multi-dimensional problems the basis functions are
tensor products of 1D basis functions. The identification of the inner product weight function
with the probability distribution of $\xxi$ simplifies the calculations of $U$'s statistical moments.
Noting that since $\Psi_0(\xxi)$ is a constant that can be normalized to satisfy 
$\left<\Psi_0,\Psi_0\right>=1$, the expectation and variance of $U$ can be computed as:
\begin{eqnarray}
 E[U]&=&\int U \, \rho(\xxi) \, \mbox{d}\xxi=\left< U,\Psi_0\right> = U_0  
 \label{eq:mean} \\
 E[(U-E[U])^2]&=&\int (U-E[U])^2 \, \rho(\xxi) \, \mbox{d}\xxi=\sum_{k=1}^P U_k^2
 \label{eq:sigma}
\left<\Psi_k,\Psi_k\right>
\end{eqnarray}

PC representations of random variables also enable efficient
approximation of the corresponding variance-based, global
sensitivity indices. This can be done by computing the so-called {\it total} 
sensitivity indices of random variables from the PC representations or Sobol decomposition~\citep{Sobol:1993,Homma:1996,Sobol:2001}. To get the total sensitivity corresponding to the uncertain
input $\xi_i$ we compute the total index~\citep{LeMaitreKnio2010,Crestaux,Sudret}:
\begin{equation} \label{equ:T-hard}
   T_i = \displaystyle \sum_{u \ni i} S_u =
         \frac{\displaystyle
              \sum_{u \ni i} \sum_{k \in K_u} U_k^2 \ave{\Psi_k^2}}
              {\displaystyle\sum_{k = 1}^P U_k^2 \ave{\Psi_k^2}} .
\end{equation}
where for each index set $u \subseteq \{1, \ldots, n\}$ we define:
\[
   K_u = \left\{ k \in \{1, \ldots, P\} :
          \begin{cases}
           \vec{\alpha}^k_i > 0, \quad \mbox{ if } i \in u,\\
           \vec{\alpha}^k_i = 0, \quad \mbox{ if } i \notin u
          \end{cases}\right\}.
        \]
        and $\vec{\alpha}^k$ is the
multi-index associated with $k^{th}$ term in the
PC expansion~\cite{LeMaitreKnio2010}.

Using Equation~\eqref{equ:T-hard}, the computation of $T_i$ is straightforward.
Note that the total sensitivity index $T_i$ measures the contribution of
the $i^{th}$ random input to total model variability by
computing the fraction of the total variance due to all the terms in the
PC expansion which involve $\xi_i$. It is also worth noting that for random variables expanded in a PC basis, all the index sets
above are dictated by the basis alone. Thus, we need to find the index sets $K_u$
for computation of $S_u$ and $\mathcal{I}_i$ for computation
of $T_i$ only once. Hence, the computation of $S_u$ or $T_i$ for a
random variable is immediate once its PC coefficients are determined.

%The series representation (\ref{eq:stochseries}) can thus be seen as
%combining approximation and probabilistic frameworks, a combination
%that has proven extremely useful in solving UQ
%problems.

\subsubsection{Non Intrusive Spectral Projection (NISP)}
The computation of the coefficients of the PC expansions $U_k$
can be done using a number of procedures. Here we adopt a non-intrusive
approach that allows the use of the forward model \emph{Geoclaw} as a black box
with no code modifications required. PC expansion coefficients are thus determined
based on a set of response \emph{Geoclaw} simulations at specified set of the uncertain parameters. 
In particular, we rely on the Non-Intrusive Spectral Projection (NISP) method that exploits the orthogonality of the basis and applies the Galerkin projection to find the PC expansion coefficients as follows:
\begin{equation}
 U_k(\bm{x},t) = \frac{\left< U, \Psi_k \right>}{\left< \Psi_k, \Psi_k \right>} = 
 \frac{1}{\left< \Psi_k, \Psi_k \right>} 
 \int U(\bm{x},t,\xxi) \Psi_k(\xxi) \rho(\xxi) \mbox{ d}\xxi
\end{equation}
This orthogonal projection minimizes the L2 error on the space spanned by the basis.
with the inner-product indicated in Equation~\ref{eq:inner}.
Using NISP the stochastic integrals are solved using a numerical quadrature to obtain:

\begin{equation}
  \left< U, \Psi_k \right> 
\approx \left< U, \Psi_k \right>_Q
= \sum_{q=1}^Q U(\xxi_q) \Psi_k(\xxi_q) \omega_q
\end{equation}
where the subscript $Q$ refers to approximating the inner product integral with
quadrature, and $\xxi_q$ and $\omega_q$ are multi-dimensional quadrature points and weights,
respectively.
The computation of the ${U}_k$ can thus be expressed as a matrix-vector product of the form:
\begin{equation} 
 U_k(\bm{x},t)=\sum_q \Pi_{kq} U(\bm{x},t,\xxi_q),\;\;\;
 \Pi_{kq}=\frac{\Psi_k(\xxi_q)\omega_q}{\left< \Psi_k, \Psi_k \right>}
\end{equation} 
where $\Pi_{kq}$ is the projection matrix and $U(\bm{x},t,\xxi_q)$ is obtained
from an ensemble of the deterministic model {\em realizations} with the uncertain parameters set at
the quadrature value $\xxi_q$.

The only cost of the NISP is that of the ensemble calculation and its storage.
The calculation of the $U(\bm{x},t,\xxi_q)$ is the most expensive part of the
inference procedure; thus, reducing the number of sampling points
while maximizing their effectiveness is critical for the procedure's
efficiency.  The quadrature order should be commensurate with the
truncation order, and should be high enough to avoid aliasing artifacts.
The choice of quadrature rule is hence critical to the performance
of the PC (in its NISP version at least). In this work, we employ the 
tensorized Gaussian quadratures as their computational costs are affordable.
